---
title: "Synthetic Medicare Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synthetic-Medicare}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CausalGPS)
library(ggplot2)
```

In this vignette, we present application of CausalGPS package on the Synthetic Medicare Data. In the dataset, we link the 2010 synthetic Medicare claims data to environmental exposures and potential confounders. The dataset is hosted on Harvard Dataverse [@khoshnevis_2022].  

## Load Data and Preprocessing

```{r}
data("synthetic_us_2010")
data <- synthetic_us_2010
knitr::kable(head((data)))
```

```{r}
# transformers
pow2 <- function(x) {x^2}
pow3 <- function(x) {x^3}
clog <- function(x) log(x+0.001)

confounders <- names(data)
confounders <- confounders[!(confounders %in% c("FIPS","Name","STATE", 
                                                "STATE_CODE","cms_mortality_pct",
                                                "qd_mean_pm25"))]
```

## Examples of Generating Pseudo Population

### Scenario 1

- Causal Inference: Matching
- GPS model: Parametric
- Optimized_compile: True

```{r}

confounders_s1 <- c("cs_poverty","cs_hispanic",
                   "cs_black",
                   "cs_ed_below_highschool",
                   "cs_median_house_value",
                   "cs_population_density",
                   "cdc_mean_bmi","cdc_pct_nvsmoker",
                   "gmet_mean_summer_tmmx",
                   "gmet_mean_summer_rmx",
                   "gmet_mean_summer_sph",
                   "cms_female_pct", "region"
)

study_data <- data[, c("qd_mean_pm25", confounders, "cms_mortality_pct")]
study_data$id <- seq_along(1:nrow(study_data))
study_data$region <- as.factor(study_data$region)
study_data$cs_PIR <- study_data$cs_median_house_value/study_data$cs_household_income

# Choose subset of data
q1 <- stats::quantile(study_data$qd_mean_pm25, 0.25)
q2 <- stats::quantile(study_data$qd_mean_pm25, 0.99)
trimmed_data <- subset(study_data[stats::complete.cases(study_data) ,],
                       qd_mean_pm25 <= q2  & qd_mean_pm25 >= q1)

trimmed_data$gmet_mean_summer_sph <- pow2(trimmed_data$gmet_mean_summer_sph)



set.seed(172)
pseudo_pop_1 <- generate_pseudo_pop(trimmed_data[, c("id", 
                                                     "cms_mortality_pct")],
                                    trimmed_data[, c("id", 
                                                     "qd_mean_pm25")],
                                    trimmed_data[, c("id", 
                                                     confounders_s1)],
                                    ci_appr = "matching",
                                    gps_density = "normal",
                                    bin_seq = NULL,
                                    exposure_trim_qtls = c(0.0 ,
                                                       1.0),
                                    use_cov_transform = TRUE,
                                    transformers = list("pow2","pow3","clog"),
                                    sl_lib = c("m_xgboost"),
                                    params = list(xgb_nrounds=c(17),
                                                  xgb_eta=c(0.28)),
                                    nthread = 1,
                                    covar_bl_method = "absolute",
                                    covar_bl_trs = 0.1,
                                    covar_bl_trs_type = "mean",
                                    max_attempt = 1,
                                    dist_measure = "l1",
                                    delta_n = 0.1,
                                    scale = 1)

plot(pseudo_pop_1)
```

### Scenario 2

- Causal Inference: Matching
- GPS model: Parametric
- Optimized_compile: False

```{r}
set.seed(172)
pseudo_pop_2 <- generate_pseudo_pop(trimmed_data[, c("id", 
                                                     "cms_mortality_pct")],
                                    trimmed_data[, c("id", 
                                                     "qd_mean_pm25")],
                                    trimmed_data[, c("id", confounders_s1)],
                                    ci_appr = "matching",
                                    gps_density = "normal",
                                    bin_seq = NULL,
                                    exposure_trim_qtls = c(0.0 ,
                                                       1.0),
                                    use_cov_transform = TRUE,
                                    transformers = list("pow2","pow3","clog"),
                                    sl_lib = c("m_xgboost"),
                                    params = list(xgb_nrounds=c(17),
                                                  xgb_eta=c(0.28)),
                                    nthread = 1,
                                    covar_bl_method = "absolute",
                                    covar_bl_trs = 0.1,
                                    covar_bl_trs_type = "mean",
                                    max_attempt = 1,
                                    dist_measure = "l1",
                                    delta_n = 0.1,
                                    scale = 1)

plot(pseudo_pop_2)

```


```{r}
optimized_data_1 <- pseudo_pop_1$pseudo_pop[,c("qd_mean_pm25","gps","counter_weight")]
nonoptimized_data_2 <- pseudo_pop_2$pseudo_pop[,c("qd_mean_pm25","gps","counter_weight")]


print(paste("Number of rows of data in the optimized approach: ",
            nrow(optimized_data_1)))

print(paste("Number of rows of data in the non-optimized approach: ",
            nrow(nonoptimized_data_2)))

print(paste("Sum of data samples in the optimized approach: ",
            sum(optimized_data_1$counter_weight)))
print(paste("Number of data in the non-optimized approach: ",
            length(nonoptimized_data_2$qd_mean_pm25)))

# Replicate gps values of optimized approach
expanded_opt_data_1 <- optimized_data_1[rep(seq_len(nrow(optimized_data_1)),
                                            optimized_data_1$counter_weight), 1:3]

exp_gps_a_1 <- expanded_opt_data_1$gps
gps_b_1 <- nonoptimized_data_2$gps

differences <- sort(gps_b_1) - sort(exp_gps_a_1)

print(paste("Sum of differences in gps values between optimized and ", 
            "non-optimized approaches is: ",
            sum(differences)))
```


### Scenario 3

- Causal Inference: Matching
- GPS model: Non-Parametric
- Optimized_compile: True


```{r}
trimmed_data <- subset(study_data[stats::complete.cases(study_data) ,],
                       qd_mean_pm25 <= q2  & qd_mean_pm25 >= q1)

set.seed(8967)
pseudo_pop_3 <- generate_pseudo_pop(trimmed_data[, c("id", 
                                                     "cms_mortality_pct")],
                                    trimmed_data[, c("id", 
                                                     "qd_mean_pm25")],
                                    trimmed_data[, c("id", confounders_s1)],
                                    ci_appr = "matching",
                                    gps_density = "kernel",
                                    bin_seq = NULL,
                                    exposure_trim_qtls = c(0.0 ,
                                                       1.0),
                                    use_cov_transform = TRUE,
                                    transformers = list("pow2","pow3","clog"),
                                    sl_lib = c("m_xgboost"),
                                    params = list(xgb_nrounds=c(12),
                                                  xgb_eta=c(0.1)),
                                    nthread = 1,
                                    covar_bl_method = "absolute",
                                    covar_bl_trs = 0.1,
                                    covar_bl_trs_type = "mean",
                                    max_attempt = 1,
                                    dist_measure = "l1",
                                    delta_n = 0.1,
                                    scale = 1)

plot(pseudo_pop_3)

```


### Scenario 4

- Causal Inference: Weighting
- GPS model: Parametric
- Optimized_compile: N/A

```{r}
trimmed_data <- subset(study_data[stats::complete.cases(study_data) ,],
                       qd_mean_pm25 <= q2  & qd_mean_pm25 >= q1)

trimmed_data$cs_poverty <- pow2(trimmed_data$cs_poverty)

set.seed(672)
pseudo_pop_4 <- generate_pseudo_pop(trimmed_data[, c("id", 
                                                     "cms_mortality_pct")],
                                    trimmed_data[, c("id", 
                                                     "qd_mean_pm25")],
                                    trimmed_data[, c("id", confounders_s1)],
                                    ci_appr = "weighting",
                                    gps_density = "normal",
                                    bin_seq = NULL,
                                    exposure_trim_qtls = c(0.0 ,
                                                       1.0),
                                    use_cov_transform = TRUE,
                                    transformers = list("pow2","pow3","clog"),
                                    sl_lib = c("m_xgboost"),
                                    params = list(xgb_nrounds=c(35),
                                                  xgb_eta=c(0.14)),
                                    nthread = 1,
                                    covar_bl_method = "absolute",
                                    covar_bl_trs = 0.1,
                                    covar_bl_trs_type = "mean",
                                    max_attempt = 1,
                                    dist_measure = "l1",
                                    delta_n = 0.1,
                                    scale = 1)

plot(pseudo_pop_4)
```

## Covariate Balance

In the previous examples, we passed specific parameters for estimating GPS values. Achieving acceptable covariate balance can be computed by searching for the most appropriate parameters and might not be a simple task. This package uses transformers on the features to get an acceptable covariate balance. The following parameters are directly related to searching for an acceptable covariate balance.

- **covar_bl_trs**: Is the acceptable threshold to stop searching. It can be computed either by `mean`, `median`, or `maximal` value of features correlation, which is defined in **covar_bl_trs_type**.
- **params**: In different iterations, we choose a parameter at random from the provided list. For example, by `xgb_nrounds=seq(1,100)` in the parameters, `nround` parameter for `xgboost` trainer will be selected a number between 1 and 100 at random, at each iteration.
- **transformers**: After each iteration, we choose a feature with the highest correlation and apply a transformer from the provided list. All transformers should be applied to a feature before reapplying the same transformer on the same feature.
- **max_attempt**: Number of test iteration. If the covar_bl_trs is not met, the search will stop after max_attempt iteration and will return the best found population. 


### Scenario 5

- Causal Inference: Matching + searching for acceptable covariate balance
- GPS model: Non-Parametric
- Optimized_compile: True

- Search domain:
  - `transformers`: `pow2`, `pow3`, `clog`.
  - `nround` for `xgboost`: 10-100.
  - `eta` for `xgboost`: 0.1-0.5.
  - `max_attempt`: 5.
  - `covar_bl_trs`: 0.08.
  - `covar_bl_trs_type`: `mean` 


```{r}
trimmed_data <- subset(study_data[stats::complete.cases(study_data) ,],
                       qd_mean_pm25 <= q2  & qd_mean_pm25 >= q1)

set.seed(328)
pseudo_pop_5 <- generate_pseudo_pop(trimmed_data[, c("id", 
                                                     "cms_mortality_pct")],
                                    trimmed_data[, c("id", 
                                                     "qd_mean_pm25")],
                                    trimmed_data[, c("id", confounders_s1)],
                                    ci_appr = "matching",
                                    gps_density = "kernel",
                                    bin_seq = NULL,
                                    exposure_trim_qtls = c(0.0 ,
                                                       1.0),
                                    use_cov_transform = TRUE,
                                    transformers = list("pow2","pow3","clog"),
                                    sl_lib = c("m_xgboost"),
                                    params = list(xgb_nrounds=seq(10, 100, 1),
                                                  xgb_eta=seq(0.1,0.5,0.01)),
                                    nthread = 1,
                                    covar_bl_method = "absolute",
                                    covar_bl_trs = 0.08,
                                    covar_bl_trs_type = "mean",
                                    max_attempt = 5,
                                    dist_measure = "l1",
                                    delta_n = 0.1,
                                    scale = 1)

plot(pseudo_pop_5)
```

In this example, after 5 attempts, we could not find a pseudo population that can satisfy the covariate balance test.


## References
